---
title: "Investigation of Central Limit Theorem using Exponential Distribution"
author: "Frank Jung"
output:
  html_document:
    css: custom.css
  pdf_document:
    fig_height: 4
    latex_engine: xelatex
---

```{r setoptions, echo=FALSE, message=FALSE, warning=FALSE}
# set our runtime environment and global defaults.
require(knitr, quietly = TRUE)
require(utils, quietly = TRUE)
require(ggplot2, quietly = TRUE)
require(scales, quietly = TRUE)
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, cache.path = "cache/", fig.path = "figure/", fig.width = 6, fig.height = 4)
options(rstudio.markdownToHTML =
    function(inputFile, outputFile) {
        require(markdown)
        markdownToHTML(inputFile, outputFile, stylesheet="custom.css")
    }
)
```

```{r simulation, echo=TRUE, message=FALSE, warning=FALSE}
set.seed(040)
lambda <- 0.2
observ <- 40
samples <- 1000
# simulated data 1000 sample rows by 40 observation columns
simdata <- matrix(sapply(1:samples, function(x) rexp(observ, lambda)), samples, observ)
# means and variances of each sample
simmeans <- apply(simdata, 1, function (x) mean(x[]))
simvars <- apply(simdata, 1, function (x) var(x[]))
# put into data frame for easy reference
simdf <- data.frame(means = simmeans, vars = simvars)
```

## Synopsis

This project will investigate the [exponential 
distribution](https://en.wikipedia.org/wiki/Exponential_distribution). Applying
the [Central Limit
Theorem](https://en.wikipedia.org/wiki/Central_limit_theorem), we will:

- show that the [sample mean converges to the theoretical mean](#sample-mean-versus-theoretical-mean) of the distribution
- compare the [sample variance to the theoretical variance](#sample-mean-versus-theoretical-mean) of the distribution
- show that the [distribution is approximately normal](#distribution-is-approximately-normal)

## Simulation

We will perform 1000 simulatations of ```r observ``` observations of the
exponential distribution with $\lambda$ of ```r lambda```.

## Sample Mean versus Theoretical Mean

Using ```r observ``` observations of the exponential distribution we will sample
it ```r samples``` times, taking the mean for each sample. The mean of these 
samples should approach the theoretical mean.

```{r samplemean, echo=FALSE, message=FALSE, warning=FALSE}
simmean <- mean(simdf$means)
ggplot(simdf, aes(x=means)) +
    geom_histogram(colour = "black", fill = "purple", alpha = 0.3) +
    geom_vline(xintercept = simmean, colour = "red", linetype = "longdash") +
    annotate("text", label = paste("sample mean", round(simmean, 2)), x = 7, y = 80, colour = "red") +
    labs(x = "sample means", y = "frequency") +
    ggtitle("Histogram of sample means from exponential distribution")
```

* The exponential distribution has a theoretical mean of $1/\lambda$ = ```r 1/lambda```.
* The sample mean after ```r samples``` draws is ```r round(mean(simdf$means), 2)```.

## Sample Variance versus Theoretical Variance

As above, take ```r observ``` observations from the exponential distribution,
this time measuring the variance. Sample these variances ```r samples``` times
and compare to the theoretical variance.

```{r samplevar, echo=FALSE, message=FALSE, warning=FALSE}
simvar <- mean(simdf$vars)
ggplot(simdf, aes(x=vars)) +
    geom_histogram(colour = "black", fill = "purple", alpha = 0.3) +
    geom_vline(xintercept = simvar, colour = "red", linetype = "longdash") +
    annotate("text", label = paste("sample variance", round(simvar, 2)), x = 55, y = 100, colour = "red") +
    labs(x = "sample variances", y = "frequency") +
    ggtitle("Histogram of sample variances from exponential distribution")
```

* The exponential distribution has a theoretical variance of $1/\lambda^2$ = ```r 1/lambda^2```.
* The sample variance after ```r samples``` draws is ```r round(mean(simdf$vars), 2)```.

## Distribution is Approximately Normal

Consider the normalised version of the independent identically distribution 
(iid) of exponential averages: $\frac {\bar X_n - \mu} {\sigma / n}$. The 
Central Limit Theorem states that this distribution tends to that of a standard 
normal as the sample size increases.
```{r uniform, echo=TRUE, message=TRUE, warning=FALSE}
set.seed(040)
# compare the distribution of 1000 random uniforms
# and distribution of 1000 averages of 40 random uniforms
means = NULL
mu = 1/lambda
se = sqrt(1/lambda^2)/samples
means <- sapply(1:samples, function(x) (mean(sample(rexp(observ, lambda), replace = TRUE)) - mu)/se)
hist(means)
# plot(rnorm(100, 0, 1), type = "l", lwd = 2)
```
