---
title: "A study of Central Limit Theorem with Exponential Distribution"
author: "Frank Jung"
output:
  pdf_document:
    toc: yes
    fig_crop: no
    keep_tex: yes
    latex_engine: xelatex
    number_sections: yes
  html_document:
    css: custom.css
graphics: yes
classoption: legalpaper
---

```{r setoptions, echo=FALSE, message=FALSE, warning=FALSE}
# set our runtime environment and global defaults.
require(knitr, quietly = TRUE)
require(ggplot2, quietly = TRUE)
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, cache.path = "cache/", fig.path = "figure/", fig.align="center")
options(rstudio.markdownToHTML =
    function(inputFile, outputFile) {
        require(markdown)
        markdownToHTML(inputFile, outputFile, stylesheet="custom.css")
    }
)
```

# Synopsis

This short project will study the [Central Limit
Theorem](https://en.wikipedia.org/wiki/Central_limit_theorem) using simulated 
calls of the [exponential 
distribution](https://en.wikipedia.org/wiki/Exponential_distribution). It will:

- show that the [sample mean converges to the theoretical mean](#sample-mean-versus-theoretical-mean) of the distribution
- compare the [sample variance to the theoretical variance](#sample-mean-versus-theoretical-mean) of the distribution
- show that the [distribution is approximately normal](#distribution-is-approximately-normal)

# Simulation

```{r simulation, echo=TRUE, message=FALSE, warning=FALSE}
set.seed(040)       # seed for random number generation
lambda <- 0.2       # exponential rate parameter
observ <- 40        # number of observations from exponential distribution
samples <- 1000     # number of samples of observations
# simulated data 1000 sample rows by 40 observation columns ...
simdata <- matrix(sapply(1:samples, function(x) rexp(observ, lambda)), samples, observ)
simmeans <- apply(simdata, 1, function (x) mean(x[]))   # mean of each sample
simvars <- apply(simdata, 1, function (x) var(x[]))     # variance of each sample
mu = 1/lambda                                           # theoretical mean
se = sqrt(1/lambda^2)/sqrt(observ)                      # theoretical standard error
simnormals <- apply(simdata, 1, function (x) (mean(x[]) - mu)/se) # normalised samples
# put into data frame for use in plots ...
simdf <- data.frame(means = simmeans, vars = simvars, normals = simnormals)
```
We will perform ```r samples``` simulatations of ```r observ``` observations 
from the exponential distribution with rate parameter ($\lambda$) of ```r lambda```.
For each of the simulations we will:

* calculate the [samples means](#sample-mean-versus-theoretical-mean)
* calculate the [samples variances](#sample-variance-versus-theoretical-variance)
* normalise the results for [later comparison](#distribution-is-approximately-normal) to the standard normal

# Sample Mean versus Theoretical Mean

The method used was to make ```r observ``` observations of the exponential
distribution repeating this ```r samples``` times. Then take the mean for each
sample of observations. The mean of these samples should approach the
theoretical mean, ($1/\lambda$), as shown in the plot below:
```{r samplemean, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(simdf, aes(x=means)) +
    geom_histogram(colour = "black", fill = "purple", alpha = 0.3) +
    geom_vline(xintercept = mean(simdf$means), colour = "red", linetype = "longdash") +
    annotate("text", label = paste("sample mean", round(mean(simdf$means), 2)), x = 7, y = 80, colour = "red") +
    labs(x = "sample means", y = "frequency") +
    ggtitle("Histogram of Sample Means from Exponential Distribution")
```

* The exponential distribution has a theoretical mean of $1/\lambda$ = ```r 1/lambda```.
* The sample mean after ```r samples``` samples is ```r round(mean(simdf$means), 2)```.

# Sample Variance versus Theoretical Variance

As before, make ```r observ``` observations of the exponential distribution 
repeating this ```r samples``` times. Then take the variance for each sample of 
observations. The mean of these sample variances should approach the 
theoretical variance, ($1/\lambda^2$) as shown in the plot below:
```{r samplevar, echo=FALSE, message=FALSE, warning=FALSE}
simvar <- mean(simdf$vars)
ggplot(simdf, aes(x=vars)) +
    geom_histogram(colour = "black", fill = "purple", alpha = 0.3) +
    geom_vline(xintercept = simvar, colour = "red", linetype = "longdash") +
    annotate("text", label = paste("sample variance", round(simvar, 2)), x = 55, y = 100, colour = "red") +
    labs(x = "sample variances", y = "frequency") +
    ggtitle("Histogram of Sample Variances from Exponential Distribution")
```

* The exponential distribution has a theoretical variance of $1/\lambda^2$ = ```r 1/lambda^2```.
* The sample variance after ```r samples``` samples is ```r round(mean(simdf$vars), 2)```.

# Distribution is Approximately Normal

If we consider the normalised version, $\frac {\bar X_n - \mu} {\sigma / n}$, of
our sample averages, then the [Central Limit 
Theorem](https://en.wikipedia.org/wiki/Central_limit_theorem) states that this 
distribution of averages tends to that of a standard normal as the sample size 
($n$) increases. Plotting a density histogram of these averages we should see a
normal like distribution:
```{r samplenormals, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data = simdf, aes(normals)) +
    geom_histogram(aes(y = ..density..), colour = "black", fill = "purple", alpha = 0.3) +
    geom_line(aes(y = ..density..), stat = "density", size = 1) +
    stat_function(fun = dnorm, colour = "red", size = 1, aes(x = c(-3, 3))) +
    labs(x = "normalised sample averages") +
    annotate("text", label = "reference: standard normal", x = 2.5, y = 0.35, colour = "red") +
    ggtitle("Density Histogram of Normalised Sample Averages")
```

To help evaluate how close this is to our simulated samples, we have overlaid
the normalised sample averages with the standard normal distribution (in red), 
$N(0,1)$.

We can further test this using the [Quantile-Quantile 
Plots](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/qqnorm.html). As
you can see the sample quantiles (black circles) closely match the standard 
normal (red line):
```{r qqnorm, echo=FALSE, message=FALSE, warning=FALSE, fig.width=5, fig.height=5}
qqnorm(simdf$normals)
qqline(simdf$normals, col = 2, lwd = 2)
```
$\hfill \square$