---
title: "Investigation of Central Limit Theorem using Exponential Distribution"
author: "Frank Jung"
date: "11 August 2015"
output: 
  html_document: 
    keep_md: yes
    toc: yes
---

## Synopsis

This project will investigate the [exponential distribution](#exponential-distribution) in R and compare it
with the [Central Limit Theorem](#central-limit-theorem). The project will:

- show that the [sample mean converges to the theoretical mean](#sample-mean-versus-theoretical-mean) of the distribution
- compare the [sample variance to the theoretical variance](#sample-mean-versus-theoretical-mean) of the distribution
- show that the [distribution is approximately normal](#distribution-is-approximately-normal)

```{r setoptions, echo=FALSE, message=FALSE, warning=FALSE}
# set our runtime environment and global defaults.
require(knitr, quietly = TRUE)
require(utils, quietly = TRUE)
require(ggplot2, quietly = TRUE)
require(scales, quietly = TRUE)
opts_chunk$set(echo = FALSE, cache = TRUE, cache.path = "cache/", 
               fig.width = 10, fig.height = 7, fig.path = "figure/")
```

## Simulation

## Sample Mean versus Theoretical Mean

## Sample Variance versus Theoretical Variance

## Distribution is Approximately Normal

## Appendices

### Central Limit Theorem

### Exponential Distribution

### Session Information

This document was produced in RStudio. The session information detailing
packages used is:

```{r sessioninformation} 
sessionInfo()
```
